Spin canonical divisors of log stacky curves

### 2015-06-26 stacky_write_up/spin_cring.tex
edited some typos

First, we deal with the case that $H^0(X,L) \neq 0$. 
Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so we must have $L = P$ where $P$ is hyperelliptic fi xed. In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_L$ is generated in degree $\leq 3$.

@@

\todo{This should be $k[x_1, \ldots, x_n]$ instead of $\sym H^0$.}
\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L,d}$ be the $d$th graded piece of $I_L$ and let $J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j}.$ For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ and $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ are independent elements.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$. By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0\\
	4 &\text{ if }\Delta > 0.
\end{cases}$$
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Chosing $g_i,h_i$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i = s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

In the case of a smooth genus 1 curve, $X$, we know $K_X = 0$, and therefore the only possibilities for half canonical divisors are $\halfcan' = 0$ or $\halfcan' = P - Q$ where $P,Q$ are distinct points, fi xed under the hyperelliptic involution. We aim to inductively construct presentations by increasing the values of the $e_i$ using Lemma ~\ref{lem:raise-stacky-order} and add additional points using Lemma ~\ref{lem:sat-1} and Lemma ~\ref{lem:sat-2}.

@@

\begin{example}
\label{eg:exception-1-5}
Let $L' = \frac{2}{5}P_1$ be a half canonical divisor on a stacky curve $\sx$. Let $u \in (P_L)_1$ be any nonzero element. We obtain $\di u|_{P_1} = 0,$ since $2P -2Q \sim 0$ and by Riemann--Roch $\dim_k (P_L)_1 = 1.$ Let $y_3 \in (R_L)_3$ be any nonzero element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally, let $y_5 \in (R_L)_5$ be an element with $y_5|_{P_1} = -P_1$. Then, we claim there is some $\alpha \in k$ so that
$R_L \cong k[u, y_3, y_5]/(y_3^4 - \alpha uy_5^2).$
The proof of this can be gotten by first using Riemann--Roch and counting the pole order at $P_1$ considerations to check the above relation exists, and that the generators and relation determine a ring whose $k$th graded piece has the same dimension of $R_L$. It is analogous to Example ~\ref{ex:base-377}, and is omittied.

@@

\ssec{Genus One Exceptional Cases}
\label{ssec:g-1-exceptional}
\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fi xed points, and $R$ be a point on $\sx$.

@@

If $\Delta > 0$, we must have $\Delta \geq 2$. In this case, let $L_X \in \di X$ be so that $2L_X \sim K_X + \Delta$. We have $\deg L_X \geq 1$, so, by Riemann--Roch, $h^0(\sx,L) \geq 1.$ Therefore, $L$ is linearly equivalent to an effective divisor, so without loss of generality, we may assume $L$ is an effective divisor. We know the theorem holds for $L' = 0$ by Example ~\ref{eg:base- 1-0}.
\todo{fi x this to accomodate the possibility that $P \in supp(L')$, this has already been done suppose ~\ref{lem:sat-1} accommodates it. }

@@

To complete the case that $\Delta > 0$, we only need show the theorem holds for $L$, assuming it holds for $L_X$. By induction, it suffices to show that if the theorem holds for a divisor $L'$ with $\deg \lfloor L' \rfloor > 0$, then it holds for $L' + \subhalf {e_i}P_i,$ with $e_i$ odd. Similarly to the above, if $P$ is not a basepoint of $L'$ then the theorem holds for $L' + \subhalf {e_i}P_i$ by emma ~\ref{lem:sat-1}, while if $P$ is a basepoint of $L'$ then the theorem holds for $L' + \subhalf {e_i}P_i$ by Lemma ~\ref{lem:sat-2}.

@@

Since $\Delta = 0$, we may write $L = L_X + \sum_{i = 1}^{r}\subhalf{e_i}P_i.$ There are now two further subcases, depending on whether $L_X = 0$ or $L_X = P - Q$ for $P$ and $Q$ two distinct hyperelliptic fi xed points.

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be even because $\halfcan_X \in \di X$ with $2\halfcan_X \sim K_X +\Delta$, where $\deg K_X$ is even). \todo{make a remark about this somewhere in the begining and reference it} In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:sat-1}. When $\delta = 0$, the proof is a rather involved induction using Lemma
~\ref{lem:sat-3} and Lemma ~\ref{lem:raise-stacky-order}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g-0-exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g-0-main}.

@@

To show that these generate all of $R'$, we need to show that all
$H^0 (\sx', d \halfcan')$ are generated by lower degrees for $d = 6$
and $7 < d \leq 24$. \todo{Should this be 25?} Note that we can again use Max Noether's theorem
(taking saturation into account) to see that

@@

\begin{center}
\textbf{Table (II)}: Genus 0 noneffective base cases generator/relations
\end{center}

@@

\begin{center}
\textbf{Table (III)}: Genus 0 noneffective base cases (Ad-iii)
\end{center}

>>>>

First, we deal with the case that $H^0(X,L) \neq 0$. 
Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so we must have $L = P$ where $P$ is hyperelliptic fixed . In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_L$ is generated in degree $\leq 3$.

@@

\todo{This should be $k[x_1, \ldots, x_n]$ instead of $\sym H^0$.}
\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L,d}$ be the $d$th graded piece of $I_L$ and let $J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j}.$ For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ and $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ are independent elements.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$. By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0\\
	4 &\text{ if }\Delta > 0.
\end{cases}$$
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Choosing $g_i,h_i$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i = s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

In the case of a smooth genus 1 curve, $X$, we know $K_X = 0$, and therefore the only possibilities for half canonical divisors are $\halfcan' = 0$ or $\halfcan' = P - Q$ where $P,Q$ are distinct points, fixed  under the hyperelliptic involution. We aim to inductively construct presentations by increasing the values of the $e_i$ using Lemma ~\ref{lem:raise-stacky-order} and add additional points using Lemma ~\ref{lem:sat-1} and Lemma ~\ref{lem:sat-2}.

@@

\begin{example}
\label{eg:exception-1-5}
Let $L' = \frac{2}{5}P_1$ be a half canonical divisor on a stacky curve $\sx$. Let $u \in (P_L)_1$ be any nonzero element. We obtain $\di u|_{P_1} = 0,$ since $2P -2Q \sim 0$ and by Riemann--Roch $\dim_k (P_L)_1 = 1.$ Let $y_3 \in (R_L)_3$ be any nonzero element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally, let $y_5 \in (R_L)_5$ be an element with $y_5|_{P_1} = -P_1$. Then, we claim there is some $\alpha \in k$ so that
$R_L \cong k[u, y_3, y_5]/(y_3^4 - \alpha uy_5^2).$
The proof of this can be gotten by first using Riemann--Roch and counting the pole order at $P_1$ considerations to check the above relation exists, and that the generators and relation determine a ring whose $k$th graded piece has the same dimension of $R_L$. It is analogous to Example ~\ref{ex:base-377}, and is omitted.

@@

\ssec{Genus One Exceptional Cases}
\label{ssec:g-1-exceptional}
\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, separably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fixed points, and $R$ be a point on $\sx$.

@@

If $\Delta > 0$, we must have $\Delta \geq 2$. In this case, let $L_X \in \di X$ be so that $2L_X \sim K_X + \Delta$. We have $\deg L_X \geq 1$, so, by Riemann--Roch, $h^0(\sx,L) \geq 1.$ Therefore, $L$ is linearly equivalent to an effective divisor, so without loss of generality, we may assume $L$ is an effective divisor. We know the theorem holds for $L' = 0$ by Example ~\ref{eg:base- 1-0}.
\todo{fix this to accommodate the possibility that $P \in supp(L')$, this has already been done suppose ~\ref{lem:sat-1} accommodates it. }

@@

To complete the case that $\Delta > 0$, we only need show the theorem holds for $L$, assuming it holds for $L_X$. By induction, it suffices to show that if the theorem holds for a divisor $L'$ with $\deg \lfloor L' \rfloor > 0$, then it holds for $L' + \subhalf {e_i}P_i,$ with $e_i$ odd. Similarly to the above, if $P$ is not a basepoint of $L'$ then the theorem holds for $L' + \subhalf {e_i}P_i$ by Lemma ~\ref{lem:sat-1}, while if $P$ is a basepoint of $L'$ then the theorem holds for $L' + \subhalf {e_i}P_i$ by Lemma ~\ref{lem:sat-2}.

@@

Since $\Delta = 0$, we may write $L = L_X + \sum_{i = 1}^{r}\subhalf{e_i}P_i.$ There are now two further subcases, depending on whether $L_X = 0$ or $L_X = P - Q$ for $P$ and $Q$ two distinct hyperelliptic fixed  points.

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be even because $\halfcan_X \in \di X$ with $2\halfcan_X \sim K_X +\Delta$, where $\deg K_X$ is even). \todo{make a remark about this somewhere in the beginning and reference it} In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:sat-1}. When $\delta = 0$, the proof is a rather involved induction using Lemma
~\ref{lem:sat-3} and Lemma ~\ref{lem:raise-stacky-order}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g-0-exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g-0-main}.

@@

To show that these generate all of $R'$, we need to show that all
$H^0 (\sx', d \halfcan')$ are generated by lower degrees for $d = 6$
and $7 < d \leq 24$. \todo{Should this be 25?} Note that we can again use Max No ether's theorem
(taking saturation into account) to see that

@@

\begin{center}
\textbf{Table (II)}: Genus 0 non-effective base cases generator/relations
\end{center}

@@

\begin{center}
\textbf{Table (III)}: Genus 0 non-effective base cases (Ad-iii)
\end{center}
